\pkg{hesim}: Health Economic Simulation Modeling and Decision Analysis

### 2021-08-14 manuscript.Rnw
Minor manuscript edits

%% -- Framework ----------------------------------------------------------------
\section{Framework} \label{sec:framework}
Economic models consist of a disease model, a utility model, and a set of cost models for each cost category. Model development proceeds in 4 steps. The first step is to set up the model by defining the model structure, target population, and treatment strategies of interest. The relevant information is then stored in a \code{hesim_data} object.

Development and analysis of the model is performed in the next three steps as shown in Figure~\ref{fig:econ-eval-process-hesim}. First, the disease progression, costs, and utility submodels are parameterized using ``estimation'' datasets. Next, the submodels are combined to construct an economic model. The economic model is then used to simulate disease progression, QALYs, and costs as a function of ``input data''. The input data always contains variables describing the target population and treatment strategies of interest, but can also contain variables related to time to facilitate use of time-varying covariates or parameters. Finally, the simulated QALYs and costs are used to perform decision analysis within a cost-effectiveness framework. While other approaches such as multi-criteria decision analysis (MCDA) could, in principle, be used as well, only CEA is currently supported. All models are simulated using PSA so that decision uncertainty can be represented.

@@

% Decision analysis
\node[draw, minimum height=4cm, minimum width=3.5cm, font=\footnotesize, text depth = 3.5cm] (C) at (5,0) {\textbf{3. Decision analysis}};
\node[draw,fill=white, minimum height=.8cm, minimum width=2.5cm, font=\footnotesize] at (5, .3){CEA};
\node[draw,fill=white, minimum height=.8cm, minimum width=2.5cm, font=\footnotesize] at (5, -.7){MCDA};

@@

This model can be simulated using any of the model types described in Section~\ref{sec:model-taxonomy}. We demonstrate use of all three below, noting that the type of model employed should depend on the available data. We begin with the richest dataset in Section~\ref{sec:ictstm-example} and show how a flexible iCTSTM can be parameterized when continuously observed patient-level multi-state data is available. The next two sections then show how models can be developed when the data is more limited. Section~\ref{sec:psm-example} uses a PSM in a case where PFS and OS are available but multi-state data is not, and Section~\ref{sec:cdtstm-example} develops a cDTSTM in a setting where multi-state data is available but it is not continuously observed. For the sake of illustration, in both Sections~\ref{sec:psm-example} and~\ref{sec:cdtstm-example}, patient-level data is available for the SOC, but not for estimation of relative treatment effects.

@@

The target population, treatment strategies, and (non-death) health states are stored in a \code{hesim_data} object and identified with integer valued identification variables (\code{patient_id}, \code{strategy_id}, and \code{state_id}). There are 3 treatment strategies: standard of care (SOC) and two new interventions (New 1 and New 2). The target population consists of a heterogeneous population of 1,000 patients, which is large enough to ensure that averages across patients in the individual simulation are reasonably stable. As we will see in the next section, patients can also be placed into subgroups (with a \code{grp_id} column) or given weights (with a \code{patient_wt} column) in the \code{patients} table; if left unspecified as done here, then all patients are assumed to belong to a single subgroup and are weighted equally.

>>>>

%% -- Framework ----------------------------------------------------------------
\section{Framework} \label{sec:framework}
Economic models in \pkg{hesim} consist of a disease model, a utility model, and a set of cost models for each cost category. Model development proceeds in 4 steps. The first step is to set up the model by defining the model structure, target population, and treatment strategies of interest. The relevant information is then stored in a \code{hesim_data} object.

Development and analysis of the model is performed in the next three steps as shown in Figure~\ref{fig:econ-eval-process-hesim}. First, the disease progression, costs, and utility submodels are parameterized using ``estimation'' datasets. Next, the submodels are combined to construct an economic model. The economic model is then used to simulate disease progression, QALYs, and costs as a function of ``input data''. The input data always contains variables describing the target population and treatment strategies of interest, but can also contain variables related to time to facilitate use of time-varying covariates or parameters. Finally, the simulated QALYs and costs are used to perform decision analysis within a cost-effectiveness framework. While other approaches such as multi-criteria decision analysis could, in principle, be used as well, only CEA is currently supported. All models are simulated using PSA so that decision uncertainty can be represented.

@@

% Decision analysis
\node[draw, minimum height=4cm, minimum width=3.5cm, font=\footnotesize, text depth = 3.5cm] (C) at (5,0) {\textbf{3. Decision analysis}};
\node[draw,fill=white, minimum height=.8cm, minimum width=2.5cm, font=\footnotesize] at (5, 0){CEA};

@@

This model can be simulated using any of the model types described in Section~\ref{sec:model-taxonomy}. We demonstrate use of all three below, noting that the type of model employed should depend on the available data. We begin with the richest dataset in Section~\ref{sec:ictstm-example} and show how a flexible iCTSTM can be parameterized when continuously observed patient-level multi-state data is available. The next two sections then show how models can be developed when the data is more limited. Section~\ref{sec:psm-example} uses a PSM in a case where PFS and OS are available but multi-state data is not, and Section~\ref{sec:cdtstm-example} develops a cDTSTM in a setting where multi-state data is available but it is not continuously observed. For the sake of illustration, in both Sections~\ref{sec:psm-example} and~\ref{sec:cdtstm-example}, patient-level data is available for the standard of care (SOC), but not for estimation of relative treatment effects.

@@

The target population, treatment strategies, and (non-death) health states are stored in a \code{hesim_data} object and identified with integer valued identification variables (\code{patient_id}, \code{strategy_id}, and \code{state_id}). There are 3 treatment strategies: SOC and two new interventions (New 1 and New 2). The target population consists of a heterogeneous population of 1,000 patients, which is large enough to ensure that averages across patients in the individual simulation are reasonably stable. As we will see in the next section, patients can also be placed into subgroups (with a \code{grp_id} column) or given weights (with a \code{patient_wt} column) in the \code{patients} table; if left unspecified as done here, then all patients are assumed to belong to a single subgroup and are weighted equally.
